### 2015-02-27 icfp15-servo/browsers.tex
Add some XXX comments.

A URL identifies a resource to load.
This resource usually consists of HTML, which is then parsed and typically turned into a Document Object
Model (DOM) tree.
From a programming languages standpoint, there are several interesting aspects of the parser design
for HTML.
First, though the specification allows the browser to abort on a parse error\footnote{\url{https://html.spec.whatwg.org/multipage/#parsing}},
in practice browsers follow the recovery algorithms described in that specification and make a best attempt to
load any page given to them.
Second, due to the presence of the \lstinline[language=HTML]{<script>} tag, the token stream can be modified
during operation.
For example, the below example that injects a close tag for the script block works in all modern browsers.
\begin{lstlisting}[language=HTML]
<html>
  <script>
    document.write (``</script>
  <li>this shows up just fine, thanks</li>
</html>
\end{lstlisting}
Finally, since resource loading is such a large factor in the latency of loading many webpages (particularly on mobile),
all modern parsers also perform token stream scanning and prefetch of resources likely to be required~\cite{browsers-slow-smartphones}.

>>>>

A URL identifies a resource to load.
This resource usually consists of HTML, which is then parsed and typically turned into a Document Object
Model (DOM) tree.
From a programming languages standpoint, there are several interesting aspects of the parser design
for HTML.
First, though the specification allows the browser to abort on a parse error\footnote{\url{https://html.spec.whatwg.org/multipage/#parsing}},
in practice browsers follow the recovery algorithms described in that specification and make a best attempt to
load any page given to them.
Second, due to the presence of the \lstinline[language=HTML]{<script>} tag, the token stream can be modified
during operation.
For example, the below example that injects a close tag for the script block works in all modern browsers.
\begin{lstlisting}[language=HTML]
<html>
  <script>
    document.write (``</script>
  <li>this shows up just fine, thanks</li>
</html>
\end{lstlisting}
% XXX what is this supposed to show? This is just a script element that contains
%     the malformed JS |document.write (``|.
Finally, since resource loading is such a large factor in the latency of loading many webpages (particularly on mobile),
all modern parsers also perform token stream scanning and prefetch of resources likely to be required~\cite{browsers-slow-smartphones}.

### 2015-02-27 icfp15-servo/servo.tex
Fix some typos and nits.
Add some XXX comments.

\subsection{Rust's syntax}
Rust provides struct and enum types (similar to Standard ML's record types and datatypes~\cite{sml97-definition}) as
well as pattern matching.
These types and associated language features provided two large benefits to Servo over traditional browsers
written in \Cplusplus{}.
First, providing new abstractions and intermediate representations is syntactically easy, so there is very little
pressure to tack additional fields into classes simply to avoid creating a large number of new header and implementation
files.
More importantly, this style of programming that uses pattern matching over concrete types instead of
virtual function calls on a class hierarchy provides a non-trivial performance gain.
Virtual functions can both have an in-memory storage cost associated with the virtual fuction tables (sometimes many thousands of bytes\footnote{\url{https://chromium.googlesource.com/chromium/blink/+/c048c5c7c2578274d82faf96e9ebda4c55e428da}}) but more importantly
incur indirect function call costs.
All C++ browser implementations transform performance-critical code to either use the \lstinline[language=C]{final}
specifier wherever possible or specialize the code in some other way to avoid this cost.

>>>>

\subsection{Rust's syntax}
% XXX Lot's of "provide" here
Rust provides struct and enum types (similar to Standard ML's record types and datatypes~\cite{sml97-definition}) as
well as pattern matching.
These types and associated language features provide two large benefits to Servo over traditional browsers
written in \Cplusplus{}.
First, providing new abstractions and intermediate representations is syntactically easy, so there is very little
pressure to tack additional fields into classes simply to avoid creating a large number of new header and implementation
files.
More importantly, this style of programming that uses pattern matching over concrete types instead of
virtual function calls on a class hierarchy provides a non-trivial performance gain.
Virtual functions can both have an in-memory storage cost associated with the virtual fuction tables (sometimes many thousands of bytes\footnote{\url{https://chromium.googlesource.com/chromium/blink/+/c048c5c7c2578274d82faf96e9ebda4c55e428da}}) but more importantly
incur indirect function call costs.
All C++ browser implementations transform performance-critical code to either use the \lstinline[language=C]{final}
specifier wherever possible or specialize the code in some other way to avoid this cost.
